\tool: Programmable, Efficient, and Scalable \\ Secure Two-Party Computation

### 2018-05-06 ccs/ld.tex
starting ccs repo
changes, comments

%\newcommand{\kw}[1]{{\lstinline[basicstyle=\small\color{blue}]{#1}}}
\newcommand{\kw}[1]{{\ensuremath{\mathtt{#1}}}}
\newcommand{\ftext}[1]{\text{\small{#1}}}
\newcommand{\cond}[3]{\ensuremath{{{#1}\:?\:{#2}\::{#3}}}}
\newcommand{\for}[4]{\ensuremath{\kw{for}\:{#1}\:\kw{in}\:[{#2}, {#3}]\:\kw{do}\:{#4}}}
\newcommand{\ite}[3]{\ensuremath{\kw{if}({#1}, {#2}, {#3})}}
\newcommand{\loops}[3]{\ensuremath{\kw{while}\:{#1} \leq {#2}\:\kw{do}\:{#3}}}
\vspace{-0.1in}
\section{Formal development}
\label{sec:ld}

@@

In this section we prove correctness and security of our \tool compiler.
%We formalize our source, intermediate and circuit languages. 
%An example of a program in our source language is provided in \figureref{ex-sml}.
%Our source runtime semantics is a model of the ideal, trusted third-party semantics, and
%generates observations corresponding to the values revealed to the
%parties.
We first formalize our source language (an example program being
\figureref{ex-sml}), and its runtime semantics.
This semantics describes the ``trusted third party"
execution semantics of the source programs and
generates observations corresponding to the values revealed to the
parties.
%
We then present the compilation rules that type check a program in the
source language and generate a program in the intermediate
language (an example program being \figureref{ex-aby}).
%
Next, we present the runtime semantics of our intermediate language
that evaluates to a circuit by ``evaluating away'' the public parts
and the arrays. Crucially, this step does not have access to the secret
inputs; those are processed by our distributed circuit semantics that
model the \mpc back-end.
%
Evaluation in this distributed setting involves the parties running an
interactive protocol. This step, like  the source semantics, emits
observations corresponding to the values revealed to the parties.
%Finally, we formalize the distributed circuit semantics (of the cryptographic back-end) that computes the generated circuit in the distributed setting, and like  the source semantics, emits observationscorresponding to the values revealed to the parties.

@@

\begin{figure}[htp]
  \footnotesize
  \[
  \begin{array}{rrcl}
    \ftext{Base type} & \sigma &::=& \kw{uint} \mid \kw{bool}\\
    \ftext{Type} & \psi &::=& \sigma \mid \sigma[n]\\
    \ftext{Constant} & c &::=& n \mid \top \mid \bot\\
    \ftext{Expression} & e &::=& c \mid x \mid e_{1} \mulop e_{2} \mid e_{1} > e_{2} \mid \cond{e_{1}}{e_{2}}{e_{3}}\\
    & &\mid& [\overline{e_{i}}]_{n} \mid x[e] \mid \kw{in}_{j}\\
    \ftext{Statement} & s &::=& \psi\:x = e \mid x := e \mid \for{x}{n_{1}}{n_{2}}{s}\\
    & & \mid& x[e_{1}] := e_{2} \mid \ite{e}{s_{1}}{s_{2}} \mid \kw{out}\:e \mid s_{1}; s_{2}\\
    & & \mid& \loops{x}{n}{s}
  \end{array}
  \]
\caption{Source language syntax}
\label{fig:srclang}
\end{figure}
\vspace{-0.1in}
\paragraph{Source language.} Our source language
 is a simple imperative language shown in \figureref{srclang}. Types
$\psi$ consist of the base types $\sigma$, and arrays
of base types $\sigma[n]$, where $n$ is the array length. Though we
model only one dimensional arrays, our implementation supports
 multi-dimensional arrays as well. Expressions $e$ in the language include the
 integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} >
e_{2}$ (we support several other
operators in the implementation, detailed in \sectionref{impl}),
conditionals $\cond{e}{e_{1}}{e_{2}}$,
array literals
$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
  similarly for other symbols) to denote a sequence of expressions.
The length of the sequence is usually clear from the context.}, and
array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
party $j$. The statements $s$ in the language comprise of variable
declarations, assignments,
\kw{for} loops, array writes, \kw{if}
statements, and sequence of statements. The statement
$\kw{out}\:e$ denotes revealing the value of $e$ to the
parties. The \kw{while} statement is an internal syntax that is not
exposed to the programmer. \\

\begin{figure}[htp]
  \footnotesize
  \fbox{$\rho \vdash e \Downarrow v$}\quad\quad
  \fbox{$\rho \vdash s \Downarrow \rho_{1}; O$}
  \[
  \\
  \begin{array}{c}
    \inferrule*[lab={\footnotesize{E-Var}}]
               {
               }
               {
                 \rho \vdash x \Downarrow \rho(x)
               }

@@

    \inferrule*[lab={\footnotesize{E-For}}]
               {
                 \rho, x \mapsto n_{1} \vdash \loops{x}{n_{2}}{s} \Downarrow \rho_{1}; O
               }
               {
                 \rho \vdash \for{x}{n_{1}}{n_{2}}{s} \Downarrow \rho_{1} - \{x\}; O
               }

@@

\end{array}
  \]
\caption{Source semantics}
\label{fig:srcsem}
\end{figure}
\vspace{-0.1in}
\paragraph{Source semantics.} The runtime semantics for the source language is shown in
\figureref{srcsem}. These semantics show how a ``trusted third party" computes the outputs when given the inputs of both the parties. 
 Values $v$, runtime environments $\rho$, and
observations $O$ are defined as follows:

@@

\newcommand{\lcond}[4]{\ensuremath{{{#2}\:?_{{#1}}\:{#3}\::{#4}}}}
%\newcommand{\for}[4]{\ensuremath{\kw{for}\:{#1}\:\kw{in}\:[{#2}, {#3}]\:\kw{do}\:{#4}}}
%\newcommand{\ite}[3]{\ensuremath{\kw{if}({#1}, {#2}, {#3})}}
%\newcommand{\loops}[3]{\ensuremath{\kw{while}\:{#1} \leq {#2}\:\kw{do}\:{#3}}}

@@

\paragraph{Intermediate language.} \figureref{interlang} shows the
intermediate language of our compiler. The syntax follows that of the source
language, except that the types and operators are \emph{labeled}. 
%Below we mainly focus on the bits that are different from the source
%language.
A label $\ell$ can be the  public label $\mathcal{P}$ or  one of the secret labels $\mathcal{A}$ or
$\mathcal{B}$, which denote arithmetic and boolean respectively.
Types $\tau$ are then
labeled base types $\sigma^{\ell}$ and arrays of labeled base types
$\sigma^{\ell}[n]$.
Most of the expression forms $\widetilde{e}$ are same as $e$, except
that the binary operators, and the conditional forms
are annotated with labels $\ell$.
Looking ahead, the label determines how the operators are
evaluated: $\mathcal{P}$-labeled operators are evaluated in-clear, $\mathcal{A}$-labeled operators generate arithmetic circuits, and
$\mathcal{B}$-labeled operators generate boolean circuits. The
form $\langle \ell \rhd m \rangle\:\widetilde{e}$ denotes coercing
$\widetilde{e}$ from label $\ell$ to label $m$.
%The statements $\widetilde{s}$ are analogous to $s$.
\vspace{-0.2in}
\paragraph{Source to intermediate compilation.} We provide the compilation
rules in \figureref{compile}. We present the rules in a
declarative style, where the rules are non-syntax
directed, and the labels $\ell$ are chosen
non-deterministically. \sectionref{impl} describes the label  inference
scheme in our implementation.

@@

     \inferrule*[lab={\footnotesize{T-For}}]
               {
                 \Gamma, x:\kw{uint}^{\mathcal{P}} \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \_
               }
               {
                 \Gamma \vdash \for{x}{n_{1}}{n_{2}}{s} \leadsto \for{x}{n_{1}}{n_{2}}{\widetilde{s}} \mid \Gamma
               }

@@

As mentioned earlier, the intermediate language models the code such
as in \figureref{ex-aby} output by our compiler.
Next, a program in the intermediate language is evaluated to a circuit
that can be executed in the distributed runtime later. The evaluation
to a circuit computes away the public parts of the program and also
\emph{flattens} the arrays so that the circuits are unaware of the
array structure.
Crucially, this phase of the semantics does not have access to the
secret inputs.
Below, we first provide the language for the circuits followed by the
evaluation rules.
%\vspace{-0.2in}
\paragraph{Evaluation to Circuits.}
\figureref{circuits} shows the syntax of circuits. A wire id range
$\wire$ denotes a set of circuit wires that carry the runtime value of a
 variable with a secret label (we will concretely define these runtime
values later as part of the circuit semantics). Circuit gates $\gate$
are wires $\wire$, input gates $\kw{in}^m_j$, multiplication gates \kw{mult}, comparison gates \kw{gt}, and multiplexer \kw{mux} gates, 
coerce gates $\langle \ell \rhd m \rangle$, and constants.
Sub-circuits $\val$ (generated from $\widetilde e$) then consist of
gates and arrays of gates.  A circuit $\crct$ is either empty, \kw{bind}-ing of a
circuit gate $\gate$ to wire $\wire$, \kw{out} gate, or a sequence of
circuits.

\begin{figure}[h]
  \footnotesize
  \fbox{$\widetilde{\rho} \vdash \widetilde{e} \Downarrow \val$}\quad\quad
  \fbox{$\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \crct$}
  \[
  \\
  \begin{array}{c}
    \inferrule*[lab={\footnotesize{S-Var}}]
               {
               }
               {
                 \widetilde{\rho} \vdash x \Downarrow \widetilde{\rho}(x)
               }

@@

Statement evaluation $\widetilde{\rho} \vdash \widetilde{s} \Downarrow
\widetilde{\rho}_{1}; \chi$ evaluates a statement $\widetilde{s}$ under 
the environment $\widetilde{\rho}$  to produce a new
environment $\widetilde{\rho}_{1}$, and a circuit $\crct$. Rules ({\sc{S-DeclC}}), ({\sc{S-DeclCkt}}), and
({\sc{S-DeclCktA}}) show the variable declaration cases. Rule
({\sc{S-DeclC}}) shows the case when $\widetilde{e}$ evaluates to
$\val$, where $\val$ is either a constant or an
array of constants. In this case, the mapping $x \mapsto
\val$ is added to the environment, and the resulting circuit
is empty. When 
$\widetilde{e}$ evaluates to a sub-circuit $\gate$, rule
({\sc{S-DeclCkt}}) picks a fresh wire $\wire$, adds the mapping $x \mapsto \wire$
to the environment $\widetilde{\rho}$, and outputs the circuit
$\kw{bind}\:\gate\:\wire$. The rule ({\sc{S-DeclCktA}}) is analogous for
$\widetilde{e}$ evaluating to an  array of sub-circuits. The variable
assignment rules (not shown in the figure) are similar. The rule
({\sc{S-WriteCkt}}) shows the case for writing to an array, where the
array contents are secret. Finally, rule ({\sc{S-Out}}) compiles to an
\kw{out} circuit.
%\vspace{-0.2in}
\paragraph{Circuit semantics.} Evaluating a program in the
intermediate language produces a
circuit to be computed using a distributed \mpc protocol. With
our circuit semantics, we model the \emph{functional} aspect
of a \mpc protocol, parametrized by cryptographic encoding and
decoding functions.

@@

\begin{figure}
  \footnotesize
  \fbox{$\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate \Downarrow b_{1}, b_{2}$}\quad\quad
  \fbox{$\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \crct \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O$}
  \[
  \\
  \begin{array}{c}
    %% \inferrule*[lab={\footnotesize{C-Wire}}]
    %%            {
    %%              b_{1}, b_{2} = \widehat{\rho_{1}}(r), \widehat{\rho_{2}}(r)
    %%            }
    %%            {
    %%              \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash r \Downarrow b_{1}, b_{2}
    %%            }               
    \inferrule*[lab={\footnotesize{C-In}}]
               {
                 (b_{1}, b_{2}) = \mathcal{E}_{m}(c)
               }
               {
                 \widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \kw{in}^{m}_{j} \Downarrow b_{1}, b_{2}
               }

@@

%\subsubsection*{Correctness theorem}
Finally, the correctness theorem is as follows (the environments on
the left of $\vdash$ are empty and we elide the environments on the
right of $\vdash$ for brevity):

@@

The theorem states that if a source statement $s$ is well-typed, and
compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source
semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
circuit semantics with the same observations $O$. 
Therefore, well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.
% We first define a judgment for
%translation of environments, $\Gamma \vdash \rho \leadsto
%\widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$. The
%judgment is defined inductively on $\Gamma$, and establishes that
%the static environment $\Gamma$ is consistent with the runtime
%environments $\rho$, $\widetilde{\rho}$, and $\widehat{\rho}_{1}$ and
%$\widehat{\rho}_{2}$. 
%As an example, if $x:\kw{uint}^{\mathcal{P}} \in
%\Gamma$, the judgment checks that $\rho(x) = \widetilde{\rho}(x) =
%n$, for some $n$. On the other hand, if $x:\kw{uint}^{m} \in
%\Gamma$ then the judgment checks that $\mathcal{D}_m( \widehat{\rho}_{1}(\widetilde{\rho}(x)),\widehat{\rho}_{2}(\widetilde{\rho}(x)))=\rho(x)$.
%%We elide the judgment, providing the full definition
%%in the supplementary material.
%With this judgment,
%the correctness theorem is as follows:
%
%\begin{theorem}[Correctness]\label{theorem:correctness}
%  $\forall\:\Gamma, s, \widetilde{s}, \Gamma_{1}, \rho, \widetilde{\rho}, \widehat{\rho}_{1}, \widehat{\rho}_{2}$, if:
%
%  \begin{enumerate}
%  \item $\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$
%  \item $\Gamma \vdash \rho \leadsto \widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$
%  \end{enumerate}
%
%  then $\exists \rho_{1}, O_{1}, \widetilde{\rho}_{1}, \chi, \widehat{\rho}'_{1}, \widehat{\rho}'_{2}, O_{2}$, s.t.
%
%  \begin{enumerate}[label=(\alph*)]
%  \item $\rho \vdash s \Downarrow \rho_{1}; O_{1}$
%  \item $\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \chi$
%  \item $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \chi \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O_{2}$
%  \item $\Gamma_{1} \vdash \rho_{1} \leadsto \widetilde{\rho}_{1}; \widehat{\rho}'_{1}, \widehat{\rho}'_{2}$
%  \item $O_{1} = O_{2}$
%  \end{enumerate}
%  
%\end{theorem}
%
%The theorem states that if a source statement $s$ is well-typed, and
%compiles to $\widetilde{s}$ in intermediate language, then $s$ terminates in the source
%semantics with observations $O_{1}$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
%circuit semantics with observations $O_{2}$, where $O_{1} = O_{2}$
%%Furthermore, the final environments are also in the compilation
%%relation. \divya{Should we call the judgment between environs as compilation relation?}
%Due to space reasons, we provide the proof in \cite{anonymized}.

>>>>

%\newcommand{\kw}[1]{{\lstinline[basicstyle=\small\color{blue}]{#1}}}
\newcommand{\kw}[1]{{\ensuremath{\mathtt{#1}}}}
\newcommand{\ftext}[1]{\text{\small{#1}}}
\newcommand{\cond}[3]{\ensuremath{{{#1}\:?\:{#2}\::{#3}}}}
\newcommand{\forl}[4]{\ensuremath{\kw{for}\:{#1}\:\kw{in}\:[{#2}, {#3}]\:\kw{do}\:{#4}}}
\newcommand{\ite}[3]{\ensuremath{\kw{if}({#1}, {#2}, {#3})}}
\newcommand{\loops}[3]{\ensuremath{\kw{while}\:{#1} \leq {#2}\:\kw{do}\:{#3}}}

\section{Formal development}
\label{sec:ld}

@@

In this section we prove correctness and security of our \tool compiler.
\dg{Put a warning that this section is technical and not prerequisite for understanding later section.}
%We formalize our source, intermediate and circuit languages. 
%An example of a program in our source language is provided in \figureref{ex-sml}.
%Our source runtime semantics is a model of the ideal, trusted third-party semantics, and
%generates observations corresponding to the values revealed to the
%parties.
We first formalize our source language (an example program being
\figureref{ex-sml}), and its runtime semantics.
This semantics describes the ``trusted third party"
execution semantics of the source programs and
generates observations corresponding to the values revealed to the
parties.
%
We then present the compilation rules that type check a program in the
source language and generate a program in the intermediate
language (an example program being \figureref{ex-aby}).
%
Next, we present the runtime semantics of our intermediate language
that evaluates to a circuit by ``evaluating away'' the public parts
and the arrays. Crucially, this step does not have access to the secret
inputs; those are processed by our distributed circuit semantics that
model the \mpc back-end.
%
Evaluation in this distributed setting involves the parties running an
interactive protocol. This step, like  the source semantics, emits
observations corresponding to the values revealed to the parties.
%Finally, we formalize the distributed circuit semantics (of the cryptographic back-end) that computes the generated circuit in the distributed setting, and like  the source semantics, emits observationscorresponding to the values revealed to the parties.

@@

\begin{figure}[htp]
  \footnotesize
  \[
  \begin{array}{rrcl}
    \ftext{Base type} & \sigma &::=& \kw{uint} \mid \kw{bool}\\
    \ftext{Type} & \psi &::=& \sigma \mid \sigma[n]\\
    \ftext{Constant} & c &::=& n \mid \top \mid \bot\\
    \ftext{Expression} & e &::=& c \mid x \mid e_{1} \mulop e_{2} \mid e_{1} > e_{2} \mid \cond{e_{1}}{e_{2}}{e_{3}}\\
    & &\mid& [\overline{e_{i}}]_{n} \mid x[e] \mid \kw{in}_{j}\\
    \ftext{Statement} & s &::=& \psi\:x = e \mid x := e \mid \forl{x}{n_{1}}{n_{2}}{s}\\
    & & \mid& x[e_{1}] := e_{2} \mid \ite{e}{s_{1}}{s_{2}} \mid \kw{out}\:e \mid s_{1}; s_{2}\\
    & & \mid& \loops{x}{n}{s}
  \end{array}
  \]
\caption{Source language syntax}
\label{fig:srclang}
\end{figure}

\paragraph{Source language.} Our source language
 is a simple imperative language shown in \figureref{srclang}. Types
$\psi$ consist of the base types $\sigma$, and arrays
of base types $\sigma[n]$, where $n$ is the array length. Though we
model only one dimensional arrays, our implementation supports
 multi-dimensional arrays as well. Expressions $e$ in the language include the
 integer constants $n$, \kw{bool} constants $\top$ and $\bot$,
variables $x$, binary operations $e_{1} \mulop e_{2}$ and $e_{1} >
e_{2}$ (we support several other
operators in the implementation, detailed in \sectionref{impl}),
conditionals $\cond{e}{e_{1}}{e_{2}}$,
array literals
$[\overline{e_{i}}]_{n}$\footnote{We write $\overline{e}$ (and
  similarly for other symbols) to denote a sequence of expressions.
The length of the sequence is usually clear from the context.}, and
array reads $x[e]$. The expression $\kw{in}_{j}$ denotes input from
party $j$. The statements $s$ in the language comprise of variable
declarations, assignments,
\kw{for} loops, array writes, \kw{if}
statements, and sequence of statements. The statement
$\kw{out}\:e$ denotes revealing the value of $e$ to the
parties\footnote{Our language also has statements $\kw{out}_1\:e$ (resp., $\kw{out}_2\:e$) to reveal value of $e$ to only the first (resp., second) party. We omit these for brevity.}. The \kw{while} statement is an internal syntax that is not
exposed to the programmer. \\

\begin{figure}[htp]
  \footnotesize
  \fbox{$\rho \vdash e \Downarrow v$}\quad\quad
  \fbox{$\rho \vdash s \Downarrow \rho_{1}; O$}
  \[
  \begin{array}{c}
    \inferrule*[lab={\footnotesize{E-Var}}]
               {
               }
               {
                 \rho \vdash x \Downarrow \rho(x)
               }

@@

    \inferrule*[lab={\footnotesize{E-For}}]
               {
                 \rho, x \mapsto n_{1} \vdash \loops{x}{n_{2}}{s} \Downarrow \rho_{1}; O
               }
               {
                 \rho \vdash \forl{x}{n_{1}}{n_{2}}{s} \Downarrow \rho_{1} - \{x\}; O
               }

@@

\end{array}
  \]
\caption{Source semantics}
\label{fig:srcsem}
\end{figure}

\paragraph{Source semantics.} The runtime semantics for the source language is shown in
\figureref{srcsem}. These semantics show how a ``trusted third party" computes the outputs when given the inputs of both the parties. 
 Values $v$, runtime environments $\rho$, and
observations $O$ are defined as follows:

@@

\newcommand{\lcond}[4]{\ensuremath{{{#2}\:?_{{#1}}\:{#3}\::{#4}}}}
%\newcommand{\forl}[4]{\ensuremath{\kw{for}\:{#1}\:\kw{in}\:[{#2}, {#3}]\:\kw{do}\:{#4}}}
%\newcommand{\ite}[3]{\ensuremath{\kw{if}({#1}, {#2}, {#3})}}
%\newcommand{\loops}[3]{\ensuremath{\kw{while}\:{#1} \leq {#2}\:\kw{do}\:{#3}}}

@@

\paragraph{Intermediate language.} \figureref{interlang} shows the
intermediate language of our compiler. The syntax follows that of the source
language, except that the types and operators are \emph{labeled}. 
%Below we mainly focus on the bits that are different from the source
%language.
A label $\ell$ can be the  public label $\mathcal{P}$ or  one of the secret labels $\mathcal{A}$ or
$\mathcal{B}$, which denote arithmetic and boolean respectively.
Types $\tau$ are then
labeled base types $\sigma^{\ell}$ and arrays of labeled base types
$\sigma^{\ell}[n]$.
Most of the expression forms $\widetilde{e}$ are same as $e$, except
that the binary operators, and the conditional forms
are annotated with labels $\ell$.
Looking ahead, the label determines how the operators are
evaluated: $\mathcal{P}$-labeled operators are evaluated in-clear, $\mathcal{A}$-labeled operators generate arithmetic circuits, and
$\mathcal{B}$-labeled operators generate boolean circuits. The
form $\langle \ell \rhd m \rangle\:\widetilde{e}$ denotes coercing
$\widetilde{e}$ from label $\ell$ to label $m$.
%The statements $\widetilde{s}$ are analogous to $s$.

\paragraph{Source to intermediate compilation.} We provide the compilation
rules in \figureref{compile}. We present the rules in a
declarative style, where the rules are non-syntax
directed, and the labels $\ell$ are chosen
non-deterministically. \sectionref{impl} describes the label  inference
scheme in our implementation.

@@

     \inferrule*[lab={\footnotesize{T-For}}]
               {
                 \Gamma, x:\kw{uint}^{\mathcal{P}} \vdash \loops{x}{n_{2}}{s} \leadsto \loops{x}{n_{2}}{\widetilde{s}} \mid \_
               }
               {
                 \Gamma \vdash \forl{x}{n_{1}}{n_{2}}{s} \leadsto \forl{x}{n_{1}}{n_{2}}{\widetilde{s}} \mid \Gamma
               }

@@

As mentioned earlier, the intermediate language models the code such
as in \figureref{ex-aby} output by our compiler.
Next, a program in the intermediate language is evaluated to a circuit
that can be executed in the distributed runtime later. The evaluation
to a circuit computes away the public parts of the program and also
\emph{flattens} the arrays so that the circuits are unaware of the
array structure.
Crucially, this phase of the semantics does not have access to the
secret inputs.
Below, we first provide the language for the circuits followed by the
evaluation rules.

\paragraph{Evaluation to Circuits.}
\figureref{circuits} shows the syntax of circuits. A wire id range
$\wire$ denotes a set of circuit wires that carry the runtime value of a
 variable with a secret label (we will concretely define these runtime
values later as part of the circuit semantics). Circuit gates $\gate$
are wires $\wire$, input gates $\kw{in}^m_j$, multiplication gates \kw{mult}, comparison gates \kw{gt}, and multiplexer \kw{mux} gates, 
coerce gates $\langle \ell \rhd m \rangle$, and constants.
Sub-circuits $\val$ (generated from $\widetilde e$) then consist of
gates and arrays of gates.  A circuit $\crct$ is either empty, \kw{bind}-ing of a
circuit gate $\gate$ to wire $\wire$, \kw{out} gate, or a sequence of
circuits.

\begin{figure}[h]
  \footnotesize
  \fbox{$\widetilde{\rho} \vdash \widetilde{e} \Downarrow \val$}\quad\quad
  \fbox{$\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \crct$}
  \[
  \begin{array}{c}
    \inferrule*[lab={\footnotesize{S-Var}}]
               {
               }
               {
                 \widetilde{\rho} \vdash x \Downarrow \widetilde{\rho}(x)
               }

@@

Statement evaluation $\widetilde{\rho} \vdash \widetilde{s} \Downarrow
\widetilde{\rho}_{1}; \chi$ evaluates a statement $\widetilde{s}$ under 
the environment $\widetilde{\rho}$  to produce a new
environment $\widetilde{\rho}_{1}$, and a circuit $\crct$. Rules ({\sc{S-DeclC}}), ({\sc{S-DeclCkt}}), and
({\sc{S-DeclCktA}}) show the variable declaration cases. Rule
({\sc{S-DeclC}}) shows the case when $\widetilde{e}$ evaluates to
$\val$, where $\val$ is either a constant or an
array of constants. In this case, the mapping $x \mapsto
\val$ is added to the environment, and the resulting circuit
is empty. When 
$\widetilde{e}$ evaluates to a sub-circuit $\gate$, rule
({\sc{S-DeclCkt}}) picks a fresh wire $\wire$, adds the mapping $x \mapsto \wire$
to the environment $\widetilde{\rho}$, and outputs the circuit
$\kw{bind}\:\gate\:\wire$. The rule ({\sc{S-DeclCktA}}) is analogous for
$\widetilde{e}$ evaluating to an  array of sub-circuits. The variable
assignment rules (not shown in the figure) are similar. The rule
({\sc{S-WriteCkt}}) shows the case for writing to an array, where the
array contents are secret. Finally, rule ({\sc{S-Out}}) compiles to an
\kw{out} circuit.

\paragraph{Circuit semantics.} Evaluating a program in the
intermediate language produces a
circuit to be computed using a distributed \mpc protocol. With
our circuit semantics, we model the \emph{functional} aspect
of a \mpc protocol, parametrized by cryptographic encoding and
decoding functions.

@@

\begin{figure}
  \footnotesize
  \fbox{$\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \gate \Downarrow b_{1}, b_{2}$}\quad\quad
  \fbox{$\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \crct \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O$}
  \[
  \begin{array}{c}
    %% \inferrule*[lab={\footnotesize{C-Wire}}]
    %%            {
    %%              b_{1}, b_{2} = \widehat{\rho_{1}}(r), \widehat{\rho_{2}}(r)
    %%            }
    %%            {
    %%              \widehat{\rho_{1}}, \widehat{\rho_{2}} \vdash r \Downarrow b_{1}, b_{2}
    %%            }               
    \inferrule*[lab={\footnotesize{C-In}}]
               {
                 (b_{1}, b_{2}) = \mathcal{E}_{m}(c)
               }
               {
                 \widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \kw{in}^{m}_{j} \Downarrow b_{1}, b_{2}
               }

@@

\paragraph{Correctness Theorem.} We prove that all well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program. That is, if a source statement $s$ is well-typed, and compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the circuit semantics with the same observations $O$. Formally, the the correctness theorem is as follows (the environments on the left of $\vdash$ are empty and we elide the environments on the right of $\vdash$ for brevity):

%\subsubsection*{Correctness theorem}
\old{
Finally, the correctness theorem is as follows (the environments on
the left of $\vdash$ are empty and we elide the environments on the
right of $\vdash$ for brevity):}

@@

\old{The theorem states that if a source statement $s$ is well-typed, and
compiles to $\widetilde{s}$ in the intermediate language, then $s$ terminates in the source
semantics with observations $O$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
circuit semantics with the same observations $O$. 
Therefore, well typed programs always terminate (array indices are always in bounds, there are no unbounded loops, etc.) and the \mpc protocol produces the same outputs as the source program.}

\dg{We prove the theorem in Appendix~\ref{}.}
% We first define a judgment for
%translation of environments, $\Gamma \vdash \rho \leadsto
%\widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$. The
%judgment is defined inductively on $\Gamma$, and establishes that
%the static environment $\Gamma$ is consistent with the runtime
%environments $\rho$, $\widetilde{\rho}$, and $\widehat{\rho}_{1}$ and
%$\widehat{\rho}_{2}$. 
%As an example, if $x:\kw{uint}^{\mathcal{P}} \in
%\Gamma$, the judgment checks that $\rho(x) = \widetilde{\rho}(x) =
%n$, for some $n$. On the other hand, if $x:\kw{uint}^{m} \in
%\Gamma$ then the judgment checks that $\mathcal{D}_m( \widehat{\rho}_{1}(\widetilde{\rho}(x)),\widehat{\rho}_{2}(\widetilde{\rho}(x)))=\rho(x)$.
%%We elide the judgment, providing the full definition
%%in the supplementary material.
%With this judgment,
%the correctness theorem is as follows:
%
%\begin{theorem}[Correctness]\label{theorem:correctness}
%  $\forall\:\Gamma, s, \widetilde{s}, \Gamma_{1}, \rho, \widetilde{\rho}, \widehat{\rho}_{1}, \widehat{\rho}_{2}$, if:
%
%  \begin{enumerate}
%  \item $\Gamma \vdash s \leadsto \widetilde{s} \mid \Gamma_{1}$
%  \item $\Gamma \vdash \rho \leadsto \widetilde{\rho}; \widehat{\rho}_{1}, \widehat{\rho}_{2}$
%  \end{enumerate}
%
%  then $\exists \rho_{1}, O_{1}, \widetilde{\rho}_{1}, \chi, \widehat{\rho}'_{1}, \widehat{\rho}'_{2}, O_{2}$, s.t.
%
%  \begin{enumerate}[label=(\alph*)]
%  \item $\rho \vdash s \Downarrow \rho_{1}; O_{1}$
%  \item $\widetilde{\rho} \vdash \widetilde{s} \Downarrow \widetilde{\rho}_{1}; \chi$
%  \item $\widehat{\rho}_{1}, \widehat{\rho}_{2} \vdash \chi \Downarrow \widehat{\rho}'_{1}, \widehat{\rho}'_{2}; O_{2}$
%  \item $\Gamma_{1} \vdash \rho_{1} \leadsto \widetilde{\rho}_{1}; \widehat{\rho}'_{1}, \widehat{\rho}'_{2}$
%  \item $O_{1} = O_{2}$
%  \end{enumerate}
%  
%\end{theorem}
%
%The theorem states that if a source statement $s$ is well-typed, and
%compiles to $\widetilde{s}$ in intermediate language, then $s$ terminates in the source
%semantics with observations $O_{1}$, $\widetilde{s}$ evaluates to circuit $\crct$, and $\crct$ terminates in the
%circuit semantics with observations $O_{2}$, where $O_{1} = O_{2}$
%%Furthermore, the final environments are also in the compilation
%%relation. \divya{Should we call the judgment between environs as compilation relation?}
%Due to space reasons, we provide the proof in \cite{anonymized}.
